% NeuroCam manual - Game Machine Handshaking
% Written by Christopher Thomas.
% Copyright (c) 2021 by Vanderbilt University. This work is released under
% the Creative Commons Attribution-ShareAlike 4.0 International License.

\chapter{Game Machine Handshaking}
\label{handshake}

The NeuroCam system queries machines on the local network to find content
providers. For the prototype system, the only network content provider is
the game machine.

The game machine should listen for UDP packets on port 8888. These will
be any of the following messages in plain text:
\begin{itemize}
\item ``\verb+looking for sources reply to port NNNN+''
\item ``\verb+talk to me on port NNNN+''
\item ``\verb+stop talking+''
\end{itemize}

The game machine may send any of the following responses:
\begin{itemize}
\item ``\verb+stream source at http://URL label XXXX+''
\item ``\verb+message source at HOST:PORT label XXXX+''
\item ``\verb+MSG (message text goes here)+''
\end{itemize}

The game machine will typically offer one video stream (the game video) and
one message source (which sends plain text timestamps for synchronization of
game events and NeuroCam data).

The video URL will generally be of the form
``\verb+http://(host IP):(port)/(file).mjpeg+''. Any valid URL should work,
as long as the file has the suffix ``\verb+mjpeg+'' and as long as the
host is given by IP address rather than hostname. This video stream will be
fetched by the NeuroCam and treated like any other camera feed.

Message sources must use IP addresses (not hostnames) as the host identifier.
These will be sent ``\verb+talk to me+'' and ``\verb+stop talking+''
messages, and when active are expected to send plain text UDP messages to
the NeuroCam machine. Messages are expected to begin with ``\verb+MSG +'',
with message content transcribed to the NeuroCam session log file. Message
source and NeuroCam timestamp information are also recorded in the log.

Multiple machines may respond to the broadcast query, and the same machine
may respond multiple times to one query. As long as the message and video
stream sources indicated by the responses are unique, they will all be
available to the NeuroCam system.

%
% This is the end of the file.
